Predicting Audio Features with Last.fm Tags

### 2023-04-06 paper/main.tex
adding state of the art

Music information retrieval (MIR) is a field that focuses on the analysis,
processing and knowledge discovery of latent information in music pieces \cite{ramirez2020machine}.

@@

% https://core.ac.uk/download/pdf/234915638.pdf

% https://sciendo.com/pdf/10.2478/ausi-2018-0009

% https://archives.ismir.net/ismir2009/paper/000018.pdf

Additionally, the Spotify audio features \dots

In general, these studies confirm the posibility of extracting knowledge from Last.fm tags.
To the best of our knowledge, no studies have addressed the problem of audio features regression, based solely on Last.fm tags.

>>>>

Music informationn retrieval (MIR) is a field that focuses on the analysis,
processing and knowledge discovery of latent information in music pieces \cite{ramirez2020machine}.

@@

In the last decade, Last.fm tags have been a popular source of metadata for MIR tasks.
Last.fm tags can contain information the genre, mood, and style of music,
and might be use to characterize certain features of a music piece.

Last.fm tags can be useful when the audio signal available,
for example, due to copyright limitations.

A number of studies have explored the use of Last.fm tags in MIR,
and have shown promising results in predicting various audio features.
% For example, tags such as "acoustic", "live", and "instrumental" have been
% shown to be good predictors of the energy and danceability of a track.
% Similarly, tags such as "happy," "sad," and "angry" have been used
% to predict the valence and arousal of a track.

Researchers have used Last.fm tags.
For example, Laurier et al. analyzed how Last.fm tags categorize mood.
In their study, they created a semantic mood space based on Last.fm tags \cite{laurier2009music}.

For example, {\c{C}}ano and Morisio discuss the process they follow to create a dataset of music
lyrics annotated with Last.fm.
In the creation process, they conclude that Last.fm tags are mostly related to music genre
and positive moods \cite{ccano2017music}.
In a similar direction, Bod{\'o} and Szil{\'a}gyi generated a dataset for lyrics genre classification
by combining the Last.fm with MusicBrainz data \cite{bodo2018connecting}.
MusicBrainz \footnotemark[1] is an online database of music editorial metadata.
\footnote[1]{https://musicbrainz.org/}

The Last.fm data has been the most widely used Last.fm dataset\footnotemark[2] in research.
This dataset is a complementary dataset of the Million Song Dataset (MSD) \cite{Bertin-Mahieux2011}.

\footnotetext[2]{Last.fm dataset, the official song tags and song similarity collection for the Million Song
Dataset, available at: http://millionsongdataset.com/lastfm.}

Additionally, the Spotify audio features have been used in multiple studies.
Historically, these features were also called "EchoNest audio features".
Echonest was an online platform was was later acquired by Spotify.
% https://en.wikipedia.org/wiki/The_Echo_Nest

Wang and Horv{\'a}t use audio features to study differences between male and female artists \cite{wang2019gender}

In general, these studies confirm the possibility of extracting knowledge from Last.fm tags.
To the best of our knowledge, no studies have addressed the problem of audio features regression, based solely on Last.fm tags.

Jamdar et al. used EchoNest audio features, combined with lyrics data to classify songs into emotion tags.
These classes were first defined based on a Last.fm tags emotion mapping \cite{jamdar2015emotion}.

Similarly, Non-negative Matrix Factorization was applied in combination with EchoNest audio features
for song recommendations \cite{benzi2016song}.

P4kxspotify is a publicly available dataset that combines music review texts with Spotify audio features.
The dataset creators argue that, although the terms of service prohibits scraping, their work is ethical \cite{pinter2020p4kxspotify}.

In general, Spotify audio features have been used as predictive variables.
We, to the best of our knowledge, are unaware of students that uses these features as target variables.

\subsubsection{Machine Learning Models}

Several studies have explored the use of machine learning
models to predict audio features from audio metadata, such as Last.fm tags.
% which studies use Last.fm tags as parameters?

The experiments conduced in this study used two classical regression models,
Boosted tree regressor and Bayesian ridge regressor.
These two machine learning models that have been widely used in MIR. % is this true? (refs)

Boosted tree regressor is a decision tree-based model that sequentially
adds weak learners to the model to improve its performance.
Bayesian ridge regressor, on the other hand,
is a probabilistic model that uses Bayesian inference to estimate
the parameters of the model.

Additionally, language models, such as GTP-2 have shown promising results in various
natural language processing (NLP) and generation.

In this particular study, GTP-2 has been fine-tuned for regression tasks,
and has shown good performance in predicting audio features from metadata.

This paper aims to apply the boosted tree regressor,
 Bayesian ridge regressor, and GTP-2 models to predict Spotify audio features from Last.fm tags.
The experiments compare the performance of these models and evaluate their effectiveness
in predicting various audio features.
The results of these experiments will provide insights into
the use of different machine learning models for MIR tasks
and can have practical applications
in music recommendation systems and genre or mood recognition.
